### 2022-04-20 outputs/paper/paper.Rmd
Data preparation and gg plot edited (2nd)

---
title: "Disagreement Regarding 2016 Presidential Election in the United States^[Code and data are available at: https://github.com/SANGWOONG-LEE/2016electionanalysis]"
subtitle: "Possible factors effecting 2016 election"
author: 
  - SangWoong Lee
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Last sentence."
output:
  bookdown::pdf_document2:
    fig_caption: yes
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(dplyr)
library(knitr)
library(scales)
```

# Introduction
First paragraph

Second paragraph

Third paragraph

This report was created using `R` [@citeR] and its libraries `tidyverse` [@tidyverse], `dplyr` [@dplyr], were used to process and clean the data, and `ggplot` [@ggplot] `bookdown` [@bookdown] and `knitr` [@knitr] was used to format this report.

# Data
```{r, include=FALSE, echo=FALSE}
# Read the reduced and cleaned data, modification done in data-preparation.r file
reduced_data1 <- read_csv("outputs/data/prepared_gss1.csv")
reduced_data2 <- read_csv("outputs/data/prepared_gss2.csv")
c_data <- read_csv("outputs/data/combined_data.csv")
```

Data Explanation

```{r fig1,echo=FALSE,fig.cap = "Distribution of national spending on foreign aid"}
c_data$nataid <- factor(c_data$nataid,                                  
                  levels = c("Too little", "About right", "Too much"))
c_data  %>%
  drop_na(nataid)  %>%
   ggplot(aes(x= nataid,group=year)) + 
    geom_bar(aes(y = ..prop.., fill = factor(..x..))) +
    geom_text(aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -0.4) +
    labs(x = element_blank(),y = "Percent", fill = "National Spending Level", title =
           "National Spending on Foreign Aid") +
    facet_grid(~year) +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(labels = c("1" = "Too little", "2" = "About right", "3" = "Too much"),
                                 values=c("#01378C","#2cd18c", "#ff3d3d"))
```

```{r fig2,echo=FALSE,fig.cap = "Distribution of national spending on arms"}
c_data$natarms <- factor(c_data$natarms,                                  
                  levels = c("Too little", "About right", "Too much"))
c_data  %>%
  drop_na(natarms)  %>%
   ggplot(aes(x= natarms,group=year)) + 
    geom_bar(aes(y = ..prop.., fill = factor(..x..))) +
    geom_text(aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -0.4) +
    labs(x = element_blank(),y = "Percent", fill = "National Spending Level", title =
           "National Spending on Arms") +
    facet_grid(~ year) +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(labels = c("1" = "Too little", "2" = "About right", "3" = "Too much"),
                                 values=c("#01378C","#2cd18c", "#ff3d3d"))
```

```{r fig3,echo=FALSE,fig.cap = "Distribution of national spending on childcare"}
c_data$natchld <- factor(c_data$natchld,                                  
                  levels = c("Too little", "About right", "Too much"))
c_data  %>%
  drop_na(natchld)  %>%
   ggplot(aes(x= natchld,group=year)) + 
    geom_bar(aes(y = ..prop.., fill = factor(..x..))) +
    geom_text(aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -0.4) +
    labs(x = element_blank(),y = "Percent", fill = "National Spending Level", title =
           "National Spending on Childcare") +
    facet_grid(~ year) +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(labels = c("1" = "Too little", "2" = "About right", "3" = "Too much"),
                                 values=c("#01378C","#2cd18c", "#ff3d3d"))
```

```{r fig4,echo=FALSE,fig.cap = "Distribution of national spending on big cities problems"}
c_data$natcity <- factor(c_data$natcity,                                  
                  levels = c("Too little", "About right", "Too much"))
c_data  %>%
  drop_na(natcity)  %>%
   ggplot(aes(x= natcity,group=year)) + 
    geom_bar(aes(y = ..prop.., fill = factor(..x..))) +
    geom_text(aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -0.4) +
    labs(x = element_blank(),y = "Percent", fill = "National Spending Level", title =
           "National Spending on Big Cities Problems") +
    facet_grid(~ year) +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(labels = c("1" = "Too little", "2" = "About right", "3" = "Too much"),
                                 values=c("#01378C","#2cd18c", "#ff3d3d"))
```

```{r fig5,echo=FALSE,fig.cap = "Distribution of national spending on halting crime rates"}
c_data$natcrime <- factor(c_data$natcrime,                                  
                  levels = c("Too little", "About right", "Too much"))
c_data  %>%
  drop_na(natcrime)  %>%
   ggplot(aes(x= natcrime,group=year)) + 
    geom_bar(aes(y = ..prop.., fill = factor(..x..))) +
    geom_text(aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -0.4) +
    labs(x = element_blank(),y = "Percent", fill = "National Spending Level", title =
           "National Spending on Halting Crime Rates") +
    facet_grid(~ year) +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(labels = c("1" = "Too little", "2" = "About right", "3" = "Too much"),
                                 values=c("#01378C","#2cd18c", "#ff3d3d"))
```

```{r fig6,echo=FALSE,fig.cap = "Distribution of national spending on dealing with drug addicts"}
c_data$natdrug <- factor(c_data$natdrug,                                  
                  levels = c("Too little", "About right", "Too much"))
c_data  %>%
  drop_na(natdrug)  %>%
   ggplot(aes(x= natdrug,group=year)) + 
    geom_bar(aes(y = ..prop.., fill = factor(..x..))) +
    geom_text(aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -0.4) +
    labs(x = element_blank(),y = "Percent", fill = "National Spending Level", title = 
           "National Spending on Dealing with Drug Addicts") +
    facet_grid(~ year) +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(labels = c("1" = "Too little", "2" = "About right", "3" = "Too much"),
                                 values=c("#01378C","#2cd18c", "#ff3d3d"))
```

```{r fig7,echo=FALSE,fig.cap = "Distribution of national spending on improving education"}
c_data$nateduc <- factor(c_data$nateduc,                                  
                  levels = c("Too little", "About right", "Too much"))
c_data  %>%
  drop_na(nateduc)  %>%
   ggplot(aes(x= nateduc,group=year)) + 
    geom_bar(aes(y = ..prop.., fill = factor(..x..))) +
    geom_text(aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -0.4) +
    labs(x = element_blank(),y = "Percent", fill = "National Spending Level", title = 
           "National Spending on Improving Education") +
    facet_grid(~ year) +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(labels = c("1" = "Too little", "2" = "About right", "3" = "Too much"),
                                 values=c("#01378C","#2cd18c", "#ff3d3d"))
```

```{r fig8,echo=FALSE,fig.cap = "Distribution of national spending on developing alternatives energy sources"}
c_data$natenrgy <- factor(c_data$natenrgy,                                  
                  levels = c("Too little", "About right", "Too much"))
c_data  %>%
  drop_na(natenrgy)  %>%
   ggplot(aes(x= natenrgy,group=year)) + 
    geom_bar(aes(y = ..prop.., fill = factor(..x..))) +
    geom_text(aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -0.4) +
    labs(x = element_blank(),y = "Percent", fill = "National Spending Level", title = 
           "National Spending on Developing Alternatives Energy") +
    facet_grid(~ year) +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(labels = c("1" = "Too little", "2" = "About right", "3" = "Too much"),
                                 values=c("#01378C","#2cd18c", "#ff3d3d"))
```

```{r fig9,echo=FALSE,fig.cap = "Distribution of national spending on protecting environment"}
c_data$natenvir <- factor(c_data$natenvir,                                  
                  levels = c("Too little", "About right", "Too much"))
c_data  %>%
  drop_na(natenvir)  %>%
   ggplot(aes(x= natenvir,group=year)) + 
    geom_bar(aes(y = ..prop.., fill = factor(..x..))) +
    geom_text(aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -0.4) +
    labs(x = element_blank(),y = "Percent", fill = "National Spending Level", title = 
           "National Spending on Protecting Environment") +
    facet_grid(~ year) +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(labels = c("1" = "Too little", "2" = "About right", "3" = "Too much"),
                                 values=c("#01378C","#2cd18c", "#ff3d3d"))
```

```{r fig10,echo=FALSE,fig.cap = "Distribution of national spending on Welfare"}
c_data$natfare <- factor(c_data$natfare,                                  
                  levels = c("Too little", "About right", "Too much"))
c_data  %>%
  drop_na(natfare)  %>%
   ggplot(aes(x= natfare,group=year)) + 
    geom_bar(aes(y = ..prop.., fill = factor(..x..))) +
    geom_text(aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -0.4) +
    labs(x = element_blank(),y = "Percent", fill = "National Spending Level", title = 
           "National Spending on Welfare") +
    facet_grid(~ year) +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(labels = c("1" = "Too little", "2" = "About right", "3" = "Too much"),
                                 values=c("#01378C","#2cd18c", "#ff3d3d"))
```

```{r fig11,echo=FALSE,fig.cap = "Distribution of national spending on improving health"}
c_data$natheal <- factor(c_data$natheal,                                  
                  levels = c("Too little", "About right", "Too much"))
c_data  %>%
  drop_na(natheal)  %>%
   ggplot(aes(x= natheal,group=year)) + 
    geom_bar(aes(y = ..prop.., fill = factor(..x..))) +
    geom_text(aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -0.4) +
    labs(x = element_blank(),y = "Percent", fill = "National Spending Level", title = 
           "National Spending on Improving Health") +
    facet_grid(~ year) +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(labels = c("1" = "Too little", "2" = "About right", "3" = "Too much"),
                                 values=c("#01378C","#2cd18c", "#ff3d3d"))
```

```{r fig12,echo=FALSE,fig.cap = "Distribution of national spending on mass transportation"}
c_data$natmass <- factor(c_data$natmass,                                  
                  levels = c("Too little", "About right", "Too much"))
c_data  %>%
  drop_na(natmass)  %>%
   ggplot(aes(x= natmass,group=year)) + 
    geom_bar(aes(y = ..prop.., fill = factor(..x..))) +
    geom_text(aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -0.4) +
    labs(x = element_blank(),y = "Percent", fill = "National Spending Level", title = 
           "National Spending on Mass Transportation") +
    facet_grid(~ year) +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(labels = c("1" = "Too little", "2" = "About right", "3" = "Too much"),
                                 values=c("#01378C","#2cd18c", "#ff3d3d"))
```

```{r fig13,echo=FALSE,fig.cap = "Distribution of national spending on park and recreation"}
c_data$natpark <- factor(c_data$natpark,                                  
                  levels = c("Too little", "About right", "Too much"))
c_data  %>%
  drop_na(natpark)  %>%
   ggplot(aes(x= natpark,group=year)) + 
    geom_bar(aes(y = ..prop.., fill = factor(..x..))) +
    geom_text(aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -0.4) +
    labs(x = element_blank(),y = "Percent", fill = "National Spending Level", title = 
           "National Spending on Park and Recreation") +
    facet_grid(~ year) +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(labels = c("1" = "Too little", "2" = "About right", "3" = "Too much"),
                                 values=c("#01378C","#2cd18c", "#ff3d3d"))
```

```{r fig14,echo=FALSE,fig.cap = "Distribution of national spending on improving the conditions of black"}
c_data$natrace <- factor(c_data$natrace,                                  
                  levels = c("Too little", "About right", "Too much"))
c_data  %>%
  drop_na(natrace)  %>%
   ggplot(aes(x= natrace,group=year)) + 
    geom_bar(aes(y = ..prop.., fill = factor(..x..))) +
    geom_text(aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -0.4) +
    labs(x = element_blank(),y = "Percent", fill = "National Spending Level", title = 
           "National Spending on Improving Conditions of Black") +
    facet_grid(~ year) +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(labels = c("1" = "Too little", "2" = "About right", "3" = "Too much"),
                                 values=c("#01378C","#2cd18c", "#ff3d3d"))
```

```{r fig15,echo=FALSE,fig.cap = "Distribution of national spending on highway and bridges"}
c_data$natroad <- factor(c_data$natroad,                                  
                  levels = c("Too little", "About right", "Too much"))
c_data  %>%
  drop_na(natroad)  %>%
   ggplot(aes(x= natroad,group=year)) + 
    geom_bar(aes(y = ..prop.., fill = factor(..x..))) +
    geom_text(aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -0.4) +
    labs(x = element_blank(),y = "Percent", fill = "National Spending Level", title = 
           "National Spending on Highway and Bridge") +
    facet_grid(~ year) +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(labels = c("1" = "Too little", "2" = "About right", "3" = "Too much"),
                                 values=c("#01378C","#2cd18c", "#ff3d3d"))
```

```{r fig16,echo=FALSE,fig.cap = "Distribution of national spending on scientific research"}
c_data$natsci <- factor(c_data$natsci,                                  
                  levels = c("Too little", "About right", "Too much"))
c_data  %>%
  drop_na(natsci)  %>%
   ggplot(aes(x= natsci,group=year)) + 
    geom_bar(aes(y = ..prop.., fill = factor(..x..))) +
    geom_text(aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -0.4) +
    labs(x = element_blank(),y = "Percent", fill = "National Spending Level", title = 
           "National Spending on Supporting Scientific Research") +
    facet_grid(~ year) +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(labels = c("1" = "Too little", "2" = "About right", "3" = "Too much"),
                                 values=c("#01378C","#2cd18c", "#ff3d3d"))
```

```{r fig17,echo=FALSE,fig.cap = "Distribution of national spending on social security"}
c_data$natsoc <- factor(c_data$natsoc,                                  
                  levels = c("Too little", "About right", "Too much"))
c_data  %>%
  drop_na(natsoc)  %>%
   ggplot(aes(x= natsoc,group=year)) + 
    geom_bar(aes(y = ..prop.., fill = factor(..x..))) +
    geom_text(aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -0.4) +
    labs(x = element_blank(),y = "Percent", fill = "National Spending Level", title = 
           "National Spending on Seocial Security") +
    facet_grid(~ year) +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(labels = c("1" = "Too little", "2" = "About right", "3" = "Too much"),
                                 values=c("#01378C","#2cd18c", "#ff3d3d"))
```

```{r fig18,echo=FALSE,fig.cap = "Respondents self Health Conditions"}
c_data$health <- factor(c_data$health,                                  
                  levels = c("Poor", "Fair", "Good", "Excellent"))
c_data  %>%
  drop_na(health)  %>%
   ggplot(aes(x= health,group=year)) + 
    geom_bar(aes(y = ..prop.., fill = factor(..x..))) +
    geom_text(aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -0.4) +
    labs(x = element_blank(),y = "Percent", fill = "Health Conditions", title = 
           "Respondents' Health Conditions by Self Assessment") +
    facet_grid(~ year) +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(labels = c("1" = "Poor", "2" = "Fair", "3" = "Good", "4" = "Excellent"),
                                 values=c("#F0E442", "#0072B2", "#D55E00", "#CC79A7"))
```

```{r fig19,echo=FALSE,fig.cap = "Government and Individuals on Health"}
c_data$helpsick <- factor(c_data$helpsick,                                  
                  levels = c("No", "Somewhat", "Both", "Yes"))
c_data  %>%
  drop_na(helpsick)  %>%
   ggplot(aes(x= helpsick,group=year)) + 
    geom_bar(aes(y = ..prop.., fill = factor(..x..))) +
    geom_text(aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -0.4) +
    labs(x = element_blank(),y = "Percent", fill = "Response", title = 
           "Should Government Be Responsible for Taking Care of Individuals?") +
    facet_grid(~ year) +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(labels = c("1" = "No", "2" = "Somewhat", "3" = "Both", "4" = "Yes"),
                                 values=c("#F0E442", "#0072B2", "#D55E00", "#CC79A7"))
```

```{r fig20,echo=FALSE,fig.cap = "Government and Individuals on Poor"}
c_data$helppoor <- factor(c_data$helppoor,                                  
                  levels = c("No", "Somewhat", "Both", "Yes"))
c_data  %>%
  drop_na(helppoor)  %>%
   ggplot(aes(x= helppoor,group=year)) + 
    geom_bar(aes(y = ..prop.., fill = factor(..x..))) +
    geom_text(aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -0.4) +
    labs(x = element_blank(),y = "Percent", fill = "Response", title = 
           "Should Government Be Responsible for Improving Standard Livings?") +
    facet_grid(~ year) +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(labels = c("1" = "No", "2" = "Somewhat", "3" = "Both", "4" = "Yes"),
                                 values=c("#F0E442", "#0072B2", "#D55E00", "#CC79A7"))

```

```{r fig21,echo=FALSE,fig.cap = "Government Reducing Income Differences"}
c_data$eqwlth <- factor(c_data$eqwlth)

c_data  %>%
  drop_na(eqwlth)  %>%
   ggplot(aes(x= eqwlth,group=year)) + 
    geom_bar(aes(y = ..prop.., fill = factor(..x..))) +
    geom_text(aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -0.4, size = 3) +
    labs(x = element_blank(),y = "Percent", fill = "Levels", title = 
           "Should Government Be Responsible for Reducing Income Differences") +
    facet_grid(~ year) +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(labels = c("1" = "Yes", "7" = "No"),
                                 values=c("#01378C","#2cd18c", "#ff3d3d","#F0E442", "#0072B2", "#D55E00", "#CC79A7"))
```

\newpage

# Results

# Discussion

## Summary

## Weaknesses and how we should proceed in the future

\newpage

\appendix

# Appendix {-}

## End Page

Thank you for participating in our survey. If you have any questions please do not hesitate to contact us.
sangwoong.lee@maill.utoronto.ca and young.suk@mail.utoronto.ca

\newpage

# References
Smith, Tom W., Davern, Michael, Freese, Jeremy, and Morgan, Stephen L., General Social Surveys, 1972-2018 [machine-readable data file] /Principal Investigator, Smith, Tom W.; Co-Principal Investigators, Michael Davern, Jeremy Freese and Stephen L. Morgan; Sponsored by National Science Foundation. --NORC ed.-- Chicago: NORC, 2019.

U.S. Census Bureau (2021). Population estimates. Retrieved from https://www.census.gov/quickfacts/fact/table/US/PST045221.